Supermassive black hole and host bulge affairs \\ II. The early and late sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram

### 2015-05-20 papers/MbhMsph/mm.tex
wed plyo

\title{Supermassive black hole and host bulge affairs \\ II. The early and late sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram.}

@@

\author{G. A. D. Savorgnan and A. W. Graham}
\affil{Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria 3122, Australia.}
\email{gsavorgn@astro.swin.edu.au}
\author{A. Marconi}
\affil{Dipartimento di Fisica e Astronomia, Universit\'a di Firenze, via G. Sansone 1, I-50019 Sesto Fiorentino, Firenze, Italy.}
\and
\author{E. Sani}
\affil{European Southern Observatory, Alonso de Cordova, Vitacura 3107, Santiago, Chile.}

@@

\begin{abstract}
In our first paper, we performed a detailed (i.e.~bulges, disks, bars, spiral arms, rings, haloes, nuclei, cores, etc.) 
decomposition of 66 galaxies, with directly measured black hole masses, 
that had been imaged at $3.6\rm~\mu m$ with \emph{Spitzer}.
Our sample is the largest to date and, for the first time, the decompositions were checked to be consistent with the galaxy kinematics.
Here we present correlations between the black hole mass, $M_{\rm BH}$, 
and the host spheroid (and galaxy) luminosity, $L_{\rm sph}$ (and $L_{\rm gal}$), 
and also stellar mass, $M_{\rm *,sph}$.
Most previous studies have used galaxy samples that were overwhelmingly dominated by high-mass, early-type objects.
Instead, our sample includes 17 spiral galaxies, half of which have $M_{\rm BH} < 10^7\rm~M_\odot$, 
and allows us to better investigate the poorly studied low-mass end of the $M_{\rm BH} - M_{\rm *,sph}$ correlation.
The bulges of early-type (E + S0) galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$, 
consistent with a dry-merging formation scenario,
and define a tight \emph{early sequence} with intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$.
On the other hand, the bulges of late-type (Sp) galaxies define a much steeper \emph{late sequence}, 
with $M_{\rm BH} \propto M_{\rm *,sph}^{2-3}$, 
indicating that gas-rich processes feed the black hole more efficiently than the host bulge as they grow. 
We additionally report that: i) S\'ersic galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.48 \pm 0.20}$, a less steep sequence than previously reported; 
ii) bulges with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
are not offset to lower $M_{\rm BH}$ from the correlation defined by bulges with $n_{\rm sph}>2$; 
and iii) $L_{\rm sph}$ and $L_{\rm gal}$ correlate equally well with $M_{\rm BH}$, in terms of intrinsic scatter, only for early-type galaxies; 
once reasonable numbers of spiral galaxies are included, the correlation with $L_{\rm sph}$ is better than that with $L_{\rm gal}$. {\bf 241 words}

@@

\cite{yee1992} also announced a relation between what was effectively black hole mass and galaxy mass for high-luminosity, bulge-dominated early-type galaxies.
This ``rough scaling'' was a premature version of the nowadays popular correlation between black hole mass, $M_{\rm BH}$,  

@@

These early studies were dominated by high-mass, early-type galaxies, 
for which they reported a quasi-linear $M_{\rm BH} - M_{\rm sph}$ relation, 
consistent with a dry-merging formation scenario which would build and maintain a linear scaling. 

@@

continued to use galaxy samples dominated by high-mass, early-type systems having $M_{\rm BH} \gtrsim 0.5 \times 10^8~\rm M_\odot$, 
and recovered a near-linear relation. 

@@

Some studies reported a slope steeper than one,  
or noticed that low-mass spheroids were offset to the lower right of the relation traced by their high-mass counterparts 

@@

\cite{graham2012bent}, \cite{grahamscott2013} and \cite{scott2013} found two distinct trends in the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - M_{\rm sph}$ diagrams, 
i.e. a linear and a super-quadratic correlation at the high- and low-mass end, respectively.
Those readers interested in an extensive review about the early discovery and successive improvements of these correlations 
should consult \cite{graham2015bulges}. \\

@@

as it does with the bulge luminosity. 
Now we roughly double the sample. \\
The $M_{\rm BH} - L_{\rm sph}$ relation for early-type (elliptical + lenticular) galaxies can be predicted from other two correlations that involve 

@@

The first of these two is the $M_{\rm BH} - \sigma$ relation \citep{ferraresemerritt2000,gebhardt2000},

@@

The second is the $L_{\rm sph} - \sigma$ relation, 

@@

The $M_{\rm BH} - L_{\rm sph}$ relation should therefore be better described by a ``broken'', rather than a single, power-law, 

@@

namely the core-S\'ersic \citep{graham2003coresersicmodel,trujillo2004coresersicmodel} and S\'ersic 

@@

whereas S\'ersic spheroids have no central deficit of stars. 
While core-S\'ersic spheroids are also ``core galaxies'', as given by the Nuker definition \citep{lauer2007lumell},
it should be noted that $\sim$20\% of ``core galaxies'' are not core-S\'ersic spheroids 
(\citealt{dullograham2014cores}, their Appendix A.2), i.e. do not have depleted cores.
The change in slope of the $L_{\rm sph} - \sigma$ relation corresponds to the division between 
core-S\'ersic and S\'ersic spheroids (e.g. \citealt{grahamguzman2003}).},

@@

In this scenario, core-S\'ersic spheroids are built in additive dry merger events, 

@@

whereas S\'ersic spheroids originate from gas-rich processes, 

@@

presented double power-law linear regressions 
for S\'ersic and core-S\'ersic spheroids in the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - M_{\rm *,sph}$ 

@@

but instead they converted $B-$band and $K_S-$band observed, total \emph{galaxy} magnitudes 

@@

These mean statistical bulge-to-total ratios were obtained from the results of literature two-component (S\'ersic-bulge/exponential-disk) decompositions. \\

@@

Several recent papers \citep{jiang2011a,jiang2013,mathur2012,reines2013} claimed an offset at the low-mass end of the $M_{\rm BH} - M_{\rm *,sph}$ diagram,
such that the black hole mass is lower than expected from the near-linear correlation traced by the high-mass, early-type spheroids. 
However, \cite{grahamscott2013} showed that the low-mass spheroids ($10^{8.5} \lesssim M_{\rm *,sph}/{\rm M_\odot} \lesssim 10^{10.5}$) 
are not randomly offset from such near-linear correlation, 
but follow the steeper relation traced by the S\'ersic spheroids. 
Using bulge/disk decomposition data from \cite{jiang2011a}, 
\cite{grahamscott2015} derived the host spheroid masses 
and showed that the steeper $M_{\rm BH} - M_{\rm *,sph}$ extends to these spheroids with even lower black hole masses 
($10^{5} \lesssim M_{\rm *,sph}/{\rm M_\odot} \lesssim 2 \times 10^{6}$), 
and a more detailed analysis will be presented in Graham et al. (2015, \emph{in preparation}).  \\

@@

for the largest sample of galaxies with directly measured black hole masses.

@@

Our decompositions were obtained from $3.6\rm~\mu m$ \emph{Spitzer} satellite imagery, 

@@

which allows us to constrain the slope of the correlation at the low-mass end.
In addition to this, our galaxy sample includes 17 spiral galaxies, 
representing a notable improvement over past studies dominated by early-type systems. 

@@

consists of 66 objects for which a dynamical measurement of the black hole mass had been reported in the literature 

@@

and for which we were able to obtain useful spheroid parameters from $3.6\rm~\mu m$ \emph{Spitzer} satellite imagery. 

@@

bulges, disks, spiral arms, bars, rings, haloes, extended or unresolved nuclear sources and partially depleted cores.

@@

to confirm the presence of rotationally supported disk components in most early-type galaxies, 

@@

(intermediate-scale disks, that are fully embedded in the bulge, 
or large-scale disks, that encase the bulge and dominate the light at large radii). 

@@

The morphological classification (E = elliptical; E/S0 = elliptical/lenticular; S0 = lenticular; S0/Sp = lenticular/spiral; Sp = spiral; and merger) 

@@

Throughout this paper, we will refer to early-type galaxies (E+S0) and late-type galaxies (Sp). 

@@

a galaxy is classified as core-S\'ersic if $\sigma > 270\rm~km~s^{-1}$,

@@

Galaxies with velocity dispersions between these bands had high-resolution images available.

@@

\section{Analysis}
\label{sec:anal}
We performed a linear regression analysis of the $M_{\rm BH} - L_{\rm gal}$ (see Table \ref{tab:lreggal}), 
$M_{\rm BH} - L_{\rm sph}$ (see Table \ref{tab:lregsph}) and $M_{\rm BH} - M_{\rm *,sph}$ (see Table \ref{tab:lregmass}) data,
using the BCES code from \cite{akritasbershady1996}. 
We also repeated the analysis using both the FITEXY routine \citep{press1992}, as modified by \cite{tremaine2002}, 
and the Bayesian estimator {\tt linmix\_err} \citep{linmixerr}. 
All of these three linear regression routines account for the intrinsic scatter, 
but only the last two allow one to quantify it.
%Because the BCES routine {\bf AL: did you wanna write (correctly) this sentence for me please? 
%can give biased results (wrong slope) when the scatter is comparable to the range of the data},
%we checked that the best-fit parameters obtained with the BCES code were consistent with those output by the modified FITEXY code.
We report linear regressions, both symmetrical and non-symmetrical, 
for S\'ersic/core-S\'ersic and for early/late-type galaxies.
Symmetrical regressions are meant to be compared with theoretical expectations, 
whereas non-symmetrical direct ($M_{\rm BH}|X$) regressions -- 
which minimize the scatter in the $\log(M_{\rm BH})$ direction -- 
can be used to predict black hole masses.

@@

\subsection{Black hole mass -- galaxy luminosity}
The $M_{\rm BH} - L_{\rm gal}$ diagram is shown in Figure \ref{fig:mbhmaggal}.
Four spiral galaxies had their total luminosities underestimated (see Section \ref{sec:data}) 
and thus are not included in the linear regression analysis (see Table \ref{tab:lreggal}). \\
\cite{lasker2014anal} analyzed a sample of 35 galaxies, among which only four were classified as spiral galaxies, 
and claimed that the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - L_{\rm gal}$ relations, 
which they fit with a single power-law, have consistent intrinsic scatter.
Here, instead, thanks to our galaxy sample that includes 17 spiral galaxies, 
we show that the claim made by \cite{lasker2014anal} is valid only for early-type galaxies. 
Our $M_{\rm BH} - L_{\rm sph}$ relation for 66 galaxies, irrespective of their morphological type, 
has an intrinsic scatter $\epsilon_{(Y|X)} = 0.51 \pm 0.06\rm~dex$ (forward linear regression) 
and $\epsilon_{(X|Y)} = 0.60 \pm 0.09\rm~dex$ (inverse linear regression), 
whereas our $M_{\rm BH} - L_{\rm gal}$ relation for 62 ($=66-4$) galaxies has $\epsilon_{(Y|X)} = 0.63 \pm 0.07\rm~dex$ 
and $\epsilon_{(X|Y)} = 0.91 \pm 0.17\rm~dex$.
However, when considering only early-type galaxies, 
we find that the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - L_{\rm gal}$ relations have the same level of intrinsic scatter. \\
%($\epsilon_{(Y|X)} = 0.41 \pm 0.06\rm~dex$ and $\epsilon_{(X|Y)} = 0.51 \pm 0.10\rm~dex$ 
%against $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$ and $\epsilon_{(X|Y)} = 0.53 \pm 0.10\rm~dex$).
Because the value of the intrinsic scatter depends on the size of the uncertainties associated with the absolute magnitudes\footnote{The 
smaller (larger) the uncertainties, the larger (smaller) the intrinsic scatter.}, 
we tested the robustness of our last conclusion by increasing such uncertainties  
(we originally assumed $0.25\rm~mag$).
The intrinsic scatter of the $M_{\rm BH} - L_{\rm gal}$ relation becomes smaller than that of the $M_{\rm BH} - L_{\rm sph}$ relation 
when assuming for $L_{\rm gal}$ an uncertainty larger than $0.7\rm~mag$, 
which would be significantly larger than the values commonly assumed in the literature, 
and would also oddly exceed the typical uncertainty that we estimated for $L_{\rm sph}$. 
Hence, we conclude that our determination of the intrinsic scatter is reliable, 
and that $M_{\rm BH}$ correlates equally well with $L_{\rm sph}$ and $L_{\rm gal}$ only for early-type galaxies, but not for all (early+late) galaxies.
%We decided to vary only the errors associated to $L_{\rm gal}$ 
%because we consider the uncertainties associated to $L_{\rm sph}$ trustworthy, since they have been estimated with an advanced method 
%that takes into account the systematic errors related to the galaxy decomposition process. 
%We found that the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - L_{\rm gal}$ relations for early-type galaxies 
%have inconsistent intrinsic scatter when assuming an error $\geq 0.7\rm~mag$ on $L_{\rm gal}$. 
%Because such error is larger than the typical error on $L_{\rm sph}$ estimated by us 
%and, in general, is significantly larger than the uncertainties on $L_{\rm gal}$ that are typically assumed in the literature, 
%we conclude that our determination of the intrinsic scatter is reliable, 
%and that $M_{\rm BH}$ correlates equally well with $L_{\rm sph}$ and $L_{\rm gal}$ only for early-type galaxies, but not for all (early+late) galaxies.
%As \cite{lasker2014anal} pointed out, from the observer's perspective, measuring galaxy luminosities is obviously easier and faster 
%than performing accurate galaxy decompositions to obtain bulge luminosities. 
%This, combined with our last conclusion, suggests that one should prefer the galaxy luminosity over the bulge luminosity 
%to predict black hole masses when dealing with samples of elliptical and lenticular galaxies only.

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{images/mbh_vs_mag_tot.pdf}
\caption{Black hole mass plotted against $3.6\rm~\mu m$ galaxy absolute magnitude. 
Symbols are coded according to the galaxy morphological type: red circle = E, red star = E/S0, 
red upward triangle = S0, blue downward triangle = S0/Sp, blue square = Sp, black diamond = merger. 
Empty symbols represent core-S\'ersic spheroids, whereas filled symbols are used for S\'ersic spheroids. 
Four spiral galaxies had their magnitudes overestimated and are shown as upper limits. 
The red dashed line indicates the BCES bisector linear regression for the 45 early-type galaxies (E+S0), 
with the red shaded area denoting its $1\sigma$ uncertainty. 
%%%We were not able to obtain any meaningful linear regression for the 13 ($=17-4$) late-type (Sp) galaxies. 
$M_{\rm BH}$ correlates equally well with $L_{\rm sph}$ and $L_{\rm gal}$ only for early-type galaxies, 
but not for all (early+late) galaxies. }
\label{fig:mbhmaggal}
\end{center}
\end{figure}

@@

and the linear regression analysis is presented in Table \ref{tab:lregsph}. \\

@@

is notably shallower than that found by GS13 ($M_{\rm BH} \propto L_{\rm sph}^{2.73 \pm 0.55}$). 

@@

the main cause of such inconsistency is that the bulge-to-total ratios obtained from our galaxy decompositions 

@@

Our sophisticated galaxy models account for the extra components, 
while the bulge-to-total ratios of GS13 were based on simple literature S\'ersic-bulge/exponential-disk decompositions 

@@

We have seen that the slope that we obtained for S\'ersic spheroids ($1.53 \pm 0.20$) is not consistent with the $2.5$ 
expected from $M_{\rm BH} \propto \sigma^5$ and $L_{\rm sph} \propto \sigma^2$, 
and also that S\'ersic and core-S\'ersic spheroids do not define two distinct $M_{\rm BH} - L_{\rm sph}$ sequences. 
We now investigate substructure in the $M_{\rm BH} - L_{\rm sph}$ diagram for early- and late-type galaxies.
First, we checked that elliptical and lenticular galaxies, taken separately, have slopes consistent with each other, 
and thus, taken together, they define a single \emph{early sequence} in the $M_{\rm BH} - L_{\rm sph}$ diagram. 
We then fit the bulges of early-type galaxies with a single log-linear regression, 

@@

We did not find any convincing evidence for the change in slope expected for consistency 
with the $M_{\rm BH} - \sigma$ and $L_{\rm sph} - \sigma$ correlations. 

@@

An alternative possibility is that there is no sharp transition going from $L_{\rm sph} \propto \sigma^2$ at low luminosities 

@@

it's mainly thanks to large integral-field-spectrograph surveys of early-type galaxies, such as the ATLAS$^{\rm 3D}$ Project \citep{cappellari2011}, 

@@

and it's now commonly accepted that most ``elliptical'' galaxies are all but simple and featureless pressure-supported systems. 
More than half of all morphologically classified elliptical galaxies in the ATLAS$^{\rm 3D}$ sample are fast rotators \citep{atlas3dIII-MNRAS}.
\cite{krajnovic2013} showed that ``fast rotators'' as a class are disk-galaxies or at least disk-like galaxies. 
In their magnitude-limited survey, systems 
without any signature of disk-like components (neither in the kinematics nor in the photometry) 
dominate only the most massive end (beyond $10^{11.5}~\rm M_\odot$) of the distribution. 
Past studies that investigated the $L_{\rm sph} - \sigma$ diagram might have failed to identify and consequently model embedded disks 
in intermediate-luminosity, early-type galaxies, thus overestimating their $L_{\rm sph}$ 
and mistakenly producing a sharp bent in the $L_{\rm sph} - \sigma$ correlation, rather than a continuously curved relation 

@@

Looking at Figure \ref{fig:mbhmagsph}, one might argue against the statistical significance of this ``tentative'' \emph{late sequence}. 
We point out the unsuitability of Pearson's and Spearman's correlation coefficients because 

>>>>

\title{Supermassive black hole and host bulge affairs \\ II. The early and late sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram}

@@

\author{Giulia A.~D.~Savorgnan and Alister W.~Graham}
\affil{Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria 3122, Australia.}
\email{gsavorgn@astro.swin.edu.au}
\author{Alessandro Marconi}
\affil{Dipartimento di Fisica e Astronomia, Universit\'a di Firenze, via G. Sansone 1, I-50019 Sesto Fiorentino, Firenze, Italy.}
\and
\author{Eleonora Sani}
\affil{European Southern Observatory, Alonso de Cordova, Vitacura 3107, Santiago, Chile.}

@@

\begin{abstract}
In our first paper, we performed a detailed (i.e.~bulge, disks, bars, spiral arms, rings, halo, nucleus, depleted core, etc.) 
decomposition of 66 galaxies, with directly measured black hole masses, 
that had been imaged at $3.6\rm~\mu m$ with \emph{Spitzer}.
Our sample is the largest to date and, for the first time, the decompositions were checked to be consistent with the galaxy kinematics.
Here we present correlations between the black hole mass, $M_{\rm BH}$, 
and the host spheroid (and galaxy) luminosity, $L_{\rm sph}$ (and $L_{\rm gal}$), 
and also stellar mass, $M_{\rm *,sph}$.
Most previous studies have used galaxy samples that were overwhelmingly dominated by high-mass, early-type galaxies.
Instead, our sample includes 17 spiral galaxies, half of which have $M_{\rm BH} < 10^7\rm~M_\odot$, 
and allows us to better investigate the poorly studied low-mass end of the $M_{\rm BH} - M_{\rm *,sph}$ correlation.
The bulges of early-type (E + S0) galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$, 
consistent with a dry-merging formation scenario,
and define a tight \emph{early-type sequence} with intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$.
At the low-mass end, the bulges of late-type (Sp) galaxies define a much steeper \emph{late-type sequence}, 
with $M_{\rm BH} \propto M_{\rm *,sph}^{2-3}$, 
indicating that gas-rich processes feed the black hole more efficiently than the host bulge as they coevolve. 
We additionally report that: i) S\'ersic galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.48 \pm 0.20}$, a less steep sequence than previously reported; 
ii) bulges with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
are not offset to lower $M_{\rm BH}$ from the correlation defined by bulges with $n_{\rm sph}>2$; 
and iii) $L_{\rm sph}$ and $L_{\rm gal}$ correlate equally well with $M_{\rm BH}$, in terms of intrinsic scatter, only for early-type galaxies; 
once reasonable numbers of spiral galaxies are included, the correlation with $L_{\rm sph}$ is better than that with $L_{\rm gal}$.

@@

\cite{yee1992} then announced a linear relation between what was effectively black hole mass and galaxy mass for high-luminosity, bulge-dominated early-type galaxies 
radiating near the Eddington limit.
This ``rough scaling'' was a premature version of the early correlations between black hole mass, $M_{\rm BH}$,  

@@

These initial studies were dominated by high-mass, early-type galaxies, 
for which they too reported a quasi-linear $M_{\rm BH} - M_{\rm sph}$ relation. 

@@

continued to use galaxy samples dominated by high-mass, early-type systems with $M_{\rm BH} \gtrsim 0.5 \times 10^8~\rm M_\odot$, 
and they too recovered a near-linear relation. 

@@

Some studies had reported a slope steeper than one,  
or noticed that the low-mass spheroids were offset to the right of (or below) the relation traced by the high-mass spheroids 

@@

\cite{graham2012bent}, \cite{grahamscott2013} and \cite{scott2013} found two distinct trends in the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - M_{\rm sph}$ diagrams:  
a linear and a super-quadratic correlation at the high- and low-mass end, respectively\footnote{Readers 
interested in an extensive review about the early discovery and successive improvements of these correlations 
should consult \cite{graham2015bulges}.}. \\

@@

as it does with the bulge luminosity.  \\
The $M_{\rm BH} - L_{\rm sph}$ relation for early-type (elliptical + lenticular) galaxies can be predicted by combining two other correlations that involve 

@@

One of these is the $M_{\rm BH} - \sigma$ relation \citep{ferraresemerritt2000,gebhardt2000},

@@

The other is the $L_{\rm sph} - \sigma$ relation, 

@@

The $M_{\rm BH} - L_{\rm sph}$ relation should therefore be better described by a ``broken'', rather than a single power-law: 

@@

namely the core-S\'ersic spheroids \citep{graham2003coresersicmodel,trujillo2004coresersicmodel} and the S\'ersic 

@@

whereas S\'ersic spheroids have no central deficit of stars.},

@@

In this scenario, core-S\'ersic spheroids are built in dry merger events 

@@

whereas S\'ersic spheroids originate from gas-rich processes 

@@

presented separate power-law linear regressions 
for the S\'ersic and core-S\'ersic spheroids in the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - M_{\rm *,sph}$ 

@@

but converted the $B-$band and $K_S-$band observed, total \emph{galaxy} magnitudes into bulge magnitudes 

@@

These mean statistical bulge-to-total ratios were obtained from the results of two-component (S\'ersic-bulge/exponential-disk) decompositions in the literature. \\

@@

%Several recent papers \citep{jiang2011a,jiang2013,mathur2012,reines2013} claimed an offset at the low-mass end of the $M_{\rm BH} - M_{\rm *,sph}$ diagram,
%such that the black hole mass is lower than expected from the near-linear correlation traced by the high-mass, early-type spheroids. 
%However, \cite{grahamscott2013} showed that the low-mass spheroids ($10^{8.5} \lesssim M_{\rm *,sph}/{\rm M_\odot} \lesssim 10^{10.5}$) 
%are not randomly offset from such near-linear correlation, 
%but follow the steeper relation traced by the S\'ersic spheroids. 
%Using bulge/disk decomposition data from \cite{jiang2011a}, 
%\cite{grahamscott2015} derived the host spheroid masses 
%and showed that the steeper $M_{\rm BH} - M_{\rm *,sph}$ extends to these spheroids with even lower black hole masses 
%($10^{5} \lesssim M_{\rm *,sph}/{\rm M_\odot} \lesssim 2 \times 10^{6}$), 
%and a more detailed analysis will be presented in Graham et al. (2015, \emph{in preparation}).  \\

@@

for galaxies with directly measured black hole masses.

@@

Our decompositions were performed on $3.6\rm~\mu m$ \emph{Spitzer} satellite imagery, 

@@

which allows us to better constrain the slope of the correlation at the low-mass end.
Furthermore, our galaxy sample includes 17 spiral galaxies, 
representing a notable improvement over past studies dominated by early-type galaxies. 

@@

consists of 66 objects for which a dynamical measurement of the black hole mass had been tabulated in the literature 

@@

and for which we were able to obtain useful spheroid parameters from $3.6\rm~\mu m$ \emph{Spitzer} satellite imagery. \\

@@

bulge, disks, spiral arms, bars, rings, halo, extended or unresolved nuclear source and partially depleted core.

@@

to confirm the presence of rotationally supported disk components in most early-type (elliptical + lenticular) galaxies, 

@@

(intermediate-scale disks that are fully embedded in the bulge, 
or large-scale disks that encase the bulge and dominate the light at large radii). 

@@

The morphological classification (E = elliptical; E/S0 = elliptical/lenticular; S0 = lenticular; S0/Sp = lenticular/spiral; Sp = spiral; and ``merger'') 

@@

Throughout this paper we will refer to early-type galaxies (E+S0) and late-type galaxies (Sp). 

@@

a spheroid is classified as core-S\'ersic if $\sigma > 270\rm~km~s^{-1}$,

@@

All of the galaxies with velocity dispersions between these two limits had high-resolution images available.

@@

\section{Analysis}
\label{sec:anal}
We performed a linear regression analysis of the $M_{\rm BH} - L_{\rm gal}$ (see Table \ref{tab:lreggal}), 
$M_{\rm BH} - L_{\rm sph}$ (see Table \ref{tab:lregsph}) and $M_{\rm BH} - M_{\rm *,sph}$ (see Table \ref{tab:lregmass}) data,
using the BCES code from \cite{akritasbershady1996}. 
We also repeated the analysis using both the FITEXY routine \citep{press1992}, as modified by \cite{tremaine2002}, 
and the Bayesian estimator {\tt linmix\_err} \citep{linmixerr}. 
All of these three linear regression routines account for the intrinsic scatter, 
but only the last two allow one to quantify it.
%Because the BCES routine {\bf AL: did you wanna write (correctly) this sentence for me please? 
%can give biased results (wrong slope) when the scatter is comparable to the range of the data},
%we checked that the best-fit parameters obtained with the BCES code were consistent with those output by the modified FITEXY code.
We report linear regressions, both symmetrical and non-symmetrical, 
for S\'ersic/core-S\'ersic and for early/late-type galaxies.
Symmetrical regressions are meant to be compared with theoretical expectations, 
whereas non-symmetrical forward ($M_{\rm BH}|X$) regressions -- 
which minimize the scatter in the $\log(M_{\rm BH})$ direction -- 
are best used to predict black hole masses.

@@

\subsection{Black hole mass -- galaxy luminosity}
The $M_{\rm BH} - L_{\rm gal}$ diagram is shown in Figure \ref{fig:mbhmaggal}.
Four spiral galaxies had their total luminosities underestimated (see Table \ref{tab:sample}) 
and thus are not included in the linear regression analysis (see Table \ref{tab:lreggal}). \\
\cite{lasker2014anal} analyzed a sample of 35 galaxies, among which only four were classified as spiral galaxies, 
and claimed that the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - L_{\rm gal}$ relations, 
which they fit with a single power-law, have consistent intrinsic scatter.
Here, instead, thanks to our galaxy sample that includes 17 spiral galaxies, 
we show that the claim made by \cite{lasker2014anal} is valid only for early-type galaxies. 
That is, when considering only early-type galaxies, 
we find that the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - L_{\rm gal}$ relations have the same level of intrinsic scatter. 
However, our $M_{\rm BH} - L_{\rm sph}$ relation for all 66 galaxies, irrespective of their morphological type, 
has an intrinsic scatter $\epsilon_{(Y|X)} = 0.51 \pm 0.06\rm~dex$ (forward linear regression) 
and $\epsilon_{(X|Y)} = 0.60 \pm 0.09\rm~dex$ (inverse linear regression), 
whereas our $M_{\rm BH} - L_{\rm gal}$ relation for 62 ($=66-4$) galaxies has $\epsilon_{(Y|X)} = 0.63 \pm 0.07\rm~dex$ 
and $\epsilon_{(X|Y)} = 0.91 \pm 0.17\rm~dex$.
%($\epsilon_{(Y|X)} = 0.41 \pm 0.06\rm~dex$ and $\epsilon_{(X|Y)} = 0.51 \pm 0.10\rm~dex$ 
%against $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$ and $\epsilon_{(X|Y)} = 0.53 \pm 0.10\rm~dex$).
Because the value of the intrinsic scatter depends on the size of the uncertainties associated with the absolute magnitudes\footnote{The 
smaller (larger) the uncertainties, the larger (smaller) the intrinsic scatter.}, 
we tested the robustness of our conclusion by increasing the uncertainties associated with the galaxy absolute magnitudes 
(we originally assumed $0.25\rm~mag$).
The intrinsic scatter of the $M_{\rm BH} - L_{\rm gal}$ relation only becomes smaller than that of the $M_{\rm BH} - L_{\rm sph}$ relation 
when assuming an uncertainty larger than $0.7\rm~mag$ for $L_{\rm gal}$, 
which would be significantly larger than the typical value commonly recognized in the literature, 
and would also oddly exceed the typical uncertainty that we estimated for $L_{\rm sph}$. 
Hence, we conclude that our determination of the relative intrinsic scatter is reliable, 
and that $M_{\rm BH}$ correlates equally well with $L_{\rm sph}$ and $L_{\rm gal}$ only for early-type galaxies, but not for all (early+late-type) galaxies.
%We decided to vary only the errors associated to $L_{\rm gal}$ 
%because we consider the uncertainties associated to $L_{\rm sph}$ trustworthy, since they have been estimated with an advanced method 
%that takes into account the systematic errors related to the galaxy decomposition process. 
%We found that the $M_{\rm BH} - L_{\rm sph}$ and $M_{\rm BH} - L_{\rm gal}$ relations for early-type galaxies 
%have inconsistent intrinsic scatter when assuming an error $\geq 0.7\rm~mag$ on $L_{\rm gal}$. 
%Because such error is larger than the typical error on $L_{\rm sph}$ estimated by us 
%and, in general, is significantly larger than the uncertainties on $L_{\rm gal}$ that are typically assumed in the literature, 
%we conclude that our determination of the intrinsic scatter is reliable, 
%and that $M_{\rm BH}$ correlates equally well with $L_{\rm sph}$ and $L_{\rm gal}$ only for early-type galaxies, but not for all (early+late) galaxies.
%As \cite{lasker2014anal} pointed out, from the observer's perspective, measuring galaxy luminosities is obviously easier and faster 
%than performing accurate galaxy decompositions to obtain bulge luminosities. 
%This, combined with our last conclusion, suggests that one should prefer the galaxy luminosity over the bulge luminosity 
%to predict black hole masses when dealing with samples of elliptical and lenticular galaxies only.

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{images/mbh_vs_mag_tot.pdf}
\caption{Black hole mass plotted against $3.6\rm~\mu m$ galaxy absolute magnitude. 
Symbols are coded according to the galaxy morphological type: red circle = E, red star = E/S0, 
red upward triangle = S0, blue downward triangle = S0/Sp, blue square = Sp, black diamond = merger. 
Empty symbols represent core-S\'ersic spheroids, whereas filled symbols are used for S\'ersic spheroids. 
Four spiral galaxies had their magnitudes overestimated (luminosities underestimated) and are shown as upper limits. 
The red dashed line indicates the BCES bisector linear regression for the 45 early-type galaxies (E+S0), 
with the red shaded area denoting its $1\sigma$ uncertainty. 
$M_{\rm BH}$ correlates equally well with $L_{\rm gal}$ and $L_{\rm sph}$ only for early-type galaxies, 
but not for all (early+late-type) galaxies. }
\label{fig:mbhmaggal}
\end{center}
\end{figure}

@@

and the linear regression analysis is presented in Table \ref{tab:lregsph}. 

@@

is notably shallower than that found by GS13 ($M_{\rm BH} \propto L_{\rm sph}^{2.73 \pm 0.55}$). \\

@@

the main cause of the inconsistency is that the bulge-to-total ratios obtained from our galaxy decompositions 

@@

Our galaxy models account for the extra components, 
while the average bulge-to-total ratios of GS13 were based on less sophisticated S\'ersic-bulge/exponential-disk decompositions 

@@

The slope that we obtained for S\'ersic spheroids ($1.53 \pm 0.20$) is not consistent with the value of $2.5$ 
expected from $M_{\rm BH} \propto \sigma^5$ and $L_{\rm sph} \propto \sigma^2$. 
In addition, the S\'ersic and core-S\'ersic spheroids appear not to define two distinct $M_{\rm BH} - L_{\rm sph}$ sequences. 
This leads us to investigate substructure in the $M_{\rm BH} - L_{\rm sph}$ diagram for early- and late-type galaxies.
First, we checked that the elliptical and lenticular galaxies, taken separately, have slopes consistent with each other, 
and thus, taken together, they define a single \emph{early-type sequence} in the $M_{\rm BH} - L_{\rm sph}$ diagram. 
We then fit the early-type galaxies with a single log-linear regression, 

@@

We did not find any convincing evidence for the change in slope required for consistency 
with the $M_{\rm BH} - \sigma$ and bent $L_{\rm sph} - \sigma$ correlations. 

@@

An additional possibility is that there is no sharp transition going from $L_{\rm sph} \propto \sigma^2$ at low luminosities 

@@

it is mainly thanks to large integral-field-spectrograph surveys of early-type galaxies, such as the ATLAS$^{\rm 3D}$ Project \citep{cappellari2011}, 

@@

and it is now commonly accepted that most ``elliptical'' galaxies contain disks. 
%More than half of all morphologically classified elliptical galaxies in the ATLAS$^{\rm 3D}$ sample are fast rotators \citep{atlas3dIII-MNRAS,krajnovic2013}.
%\cite{krajnovic2013} showed that ``fast rotators'' as a class are disk-galaxies or at least disk-like galaxies. 
%In their magnitude-limited survey, systems 
%without any signature of disk-like components (neither in the kinematics nor in the photometry) 
%dominate only the most massive end (beyond $10^{11.5}~\rm M_\odot$) of the distribution. 
Past studies that investigated the $L_{\rm sph} - \sigma$ diagram might have failed to identify and consequently model the disks 
in intermediate-luminosity, early-type galaxies, thus overestimating $L_{\rm sph}$ 
and mistakenly producing a sharp bend in the $L_{\rm sph} - \sigma$ correlation, rather than a continuously curved relation 

@@

From a cursory inspection of Figure \ref{fig:mbhmagsph}, one might be tempted to doubt the statistical significance of this ``tentative'' \emph{late-type sequence}. 
However, a visual inspection of the plotted data requires one to take into account the error bars 
Similarly, the Pearson's and Spearman's correlation coefficients are not applicable because 
